### 2019-02-22 manuscript/introduction.tex
revision of SI intro paragraph

% Outline, discussed by Boris and Emma:
%
% we are interested in the evolution of traits and their influence on lineage diversification
%     ref Barrett work 
% 		no diversification?, but trait shifts:
%		barrett_2008: "Because polyploidy affects the entire genome, it is perhaps not surprising that it influences many aspects of the phenotype, including the mating system. However, although it has long been recognized that the evolutionary transition from diploidy to polyploidy may result in correlated changes in mating patterns, the theoretical and empirical evidence is limited and often contradictory." (p.5)
%     any one study focuses on one/few traits
%     but any real organisms have many, so need to enlarge our view
%     mini "here, we..."
% we have tools to try to answer these questions
%     sse models
%     issues
%     but a way forward is to allow for multiple factors that could influence diversification
% ploidy
%     Stebbins
%     Mayrose, Soltis
%     recent paleo WGD, diploidization
% breeding system
%     directionality
%     diversification in Solanaceae
%     correlation with ploidy
%     pathways from SI-D to SC-P: direct or via SC-D
% here, we...
%     diversification: interaction between ploidy and breeding system
%     also: diploidization, pathways

@@

% E plan, merging the text from R and B:
%
% polyploidy is potentially important, widespread
% link with diversification; include diploidization hypothesis
%
% link with other traits -- correlated (geography), more directly connected (selfing, si)
%
% si
% si and ploidy
% pathways

@@

From the very beginning of polyploidy research, the study of this cytogenetic property considered correlations with other traits \citep{stebbins1938}. 
Some of these associations are driven by indirect or ecological effects, such as the tendency for polyploids, and other traits, to be found at higher latitudes or more marginal habitats. % todo refs: husband etc
Other correlations may have a more direct, causal connection.
Among the many of changes associated with polyploidization, perhaps the most prominent is the association between polyploidy and propensity for self-fertilization \citep{stebbins1950, barrett1988}.
In some cases, the evidence for a correlated shift in mating system along with polyploidization appears limited and sometimes contradictory \citep{barringer2007, barrett2008, husband2008}.
In other cases, however, polyploidy is not only a suspected correlate of breeding systems but indeed a causal link \citep{stout1942, lewis1947}.
Doubled number of alleles in pollen is thought to effect disruption of the genetic mechanisms in gametophytic self-incompatibility systems, which prevent self-fertilization \citep{entani1999, tsukamoto2005, kubo2010}. 
In clades with these systems, it is thus natural to consider the simultanous macroevolution of ploidy and breeding system.

% E note to self: edit this next...
Breeding system shifts---changes in the collection of physiological and morphological traits that determine the likelihood that any two gametes unite---are remarkably common and affect the distribution and amount of genetic variation in populations \citep{stebbins1974,barrett2013}.
They are by themselves thought to have a broad effect on other traits. 
The frequent transition from self-incompatibility (SI) to self-compatibility (SC) is strongly associated with net diversification rate changes \citep{goldberg_2010,devos2014}.
Given that changes in ploidy and breeding systems may be causally related and have profound affects on the fate of lineages, it seems particularly profitable to examine whether and how their macroevolutionary effects interact.
Moreover, increases in ploidy are known to directly affect breeding system states and present a rare opportunity for the study of paths and path-dependence on trait evolution and diversification.
For example, we could ask if the ordinary mutation (and fixation) causes more losses of SI than polyploidization, whether more SC polyploids are generated directly, after polyploidization of diploid SI lineages, or through a two-step pathway, involving loss of SC and subsequent polyploidization of a SC lineage.
And, are diversification rates of polyploid SC lineages different if one or other path was taken? \citet{charlesworth1985} 
%
%trans-polymorphism: We also have this great data for SI loci and the transition could be even more common, providing power to resolve the great diversification questions
%
% set up pathway dependence from SI-D to SC-P: direct or via SC-D

Here, we [...]
% Someone else should do this, once we have the full data.
% diversification: interaction between ploidy and breeding system
% 	examine individual effects
% 	examine joint effects
% 	with and without HMMs
% 	allow for re-diploidization or not
% also: diploidization, pathways
% 	examine pathways from SI-D to SC-P: direct or via SC-D

%The second question (talk here about breeding system and polyploidy, and how this could be different across clades and this is one of the reasons why we focused on Solanaceae).
%-Diversification and breeding systems\\
%Goldberg and Igic 2012, different  perspectives\\

>>>>

@@

@@

From the very beginning of polyploidy research, the study of this cytogenetic property considered correlations with other traits \citep{stebbins1938}. 
Some of these associations are driven by indirect or ecological effects, such as the tendency for polyploidy, and other traits, to be found at higher latitudes or marginal habitats. % todo refs: husband etc
Other correlations may have a more direct, causal connection.
Among the many changes associated with polyploidization, perhaps the most prominent is the association between polyploidy and propensity for self-fertilization \citep{stebbins1950, barrett1988}.
In some cases, the evidence for a correlated shift in mating system along with polyploidization appears limited and sometimes contradictory \citep{barringer2007, barrett2008, husband2008}.
% barrett2008: "Because polyploidy affects the entire genome, it is perhaps not surprising that it influences many aspects of the phenotype, including the mating system. However, although it has long been recognized that the evolutionary transition from diploidy to polyploidy may result in correlated changes in mating patterns, the theoretical and empirical evidence is limited and often contradictory." (p.5)
In other cases, however, polyploidy is not only a suspected correlate of breeding systems but indeed a causal link \citep{stout1942, lewis1947}.
Doubled number of alleles in pollen is thought to effect disruption of the genetic mechanisms in gametophytic self-incompatibility systems, which prevent self-fertilization \citep{entani1999, tsukamoto2005, kubo2010}. 
This creates a correlation between polyploidy and self-compatibility by preventing the existence of self-incompatible polyploids.
In clades with these systems, it is thus natural to consider the simultanous macroevolution of ploidy and breeding system.

% E note to self: edit this next...
Breeding system shifts---changes in the collection of physiological and morphological traits that determine the likelihood that any two gametes unite---are remarkably common and affect the distribution and amount of genetic variation in populations \citep{stebbins1974, barrett2013}.
In particular, self-incompatibility (SI) systems cause a plant to reject its own pollen, and their loss, yielding self-compatibility (SC), is one of the most replicated transitions in flowering plant evolution. % refs: stebbins, igic
Previous phylogenetic analyses have reported higher rates of diversification for SI than for SC lineages \citep{goldberg_2010, devos2014}, but they have not considered the possibility of other correlated traits driving this pattern.
Given that changes in ploidy and breeding systems may be causally related and have profound affects on the fate of lineages, it seems particularly profitable to examine possible interactions in their macroevolutionary effects.
This includes their joint influence on lineage diversification, and also potential patterns in the order of their transitions.
For example, do losses of SI more commonly occur tied to polyploidization, or without a ploidy shift?
Do polyploids arise more commonly from SI or SC diploids?
\Citet{robertson_2011} found that the pathway from SI diploids to SC polyploids is dominated by loss of SI followed later by polyploidization over long timescales, but proceeds in one step via polyploidization of SI species over short timescales.
We revisit this question with a greatly improved phylogeny and methods that allow for diversification rate differences. % E: Technically, we allow for div effects in the rate estimates, but not in the pathways.  If SC-D have so much extinction that the step to SC-P never happens, the old pathway methods won't see that.  I'll think about if this is easy to fix.
% And, are diversification rates of polyploid SC lineages different if one or other path was taken? \citet{charlesworth1985} % E: This is a great question.  We don't answer it with this round of analyses.  Maybe when we have more time for the revision?

% note: saving trans-specific polymorphism for Methods
% B: trans-polymorphism: We also have this great data for SI loci and the transition could be even more common, providing power to resolve the great diversification questions

% here, we...
%     diversification: interaction between ploidy and breeding system
%     also: diploidization, pathways
Here, we [...]
% Someone else should do this, once we have the full data.
% diversification: interaction between ploidy and breeding system
% 	examine individual effects
% 	examine joint effects
% 	with and without HMMs
% 	allow for re-diploidization or not
% also: diploidization, pathways
% 	examine pathways from SI-D to SC-P: direct or via SC-D
